{\bf {\Large Detiding time series of dissolved oxygen for improved estimates of ecosystem metabolism}}

### 2014-08-25 swmp_noise.tex
complete first draft

\usepackage{alltt}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{indentfirst}
\usepackage{titlesec}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{outlines}
\usepackage{lineno}
\usepackage{array}
\usepackage{times}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage[position=t]{subfig}
\usepackage{paralist}
\usepackage[noae]{Sweave}
\usepackage{natbib}
\usepackage{array}
\usepackage{pdflscape}
\usepackage{bm}
\usepackage{showlabels}
\bibpunct{(}{)}{,}{a}{}{,}

@@

Multiple time series were simulated by varying the conditions in \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}. Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, and degree of association of the tide with the \ac{DO} signal.  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$. A total of 81 time series were created based on the unique combinations of parameters (\cref{fig:sim_ex}).  Three half window widths for each variable (day, hour of day, and tide height) were evaluated: decimal time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.  The window widths were chosen based on preliminary assessments that suggested variation in model performance was adequately captured by these values.  In total, 27 window width combinations were evaluated for each of 81 simulated time series, producing results for 2187 weighted regressions.

The detided or normalized values for each regression were compared to the simulated data to evaluate the ability of weighted regression to reproduce the biological \ac{DO} signals. Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the detided time series and the biological \ac{DO} time series as a known component of the observed.  Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).  Results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section.  Mean correlation for all time series and window widths between the detided and biological values was 0.60, with values ranging from -0.78 to 1.00.  Mean error was 1.21, with values ranging from 0.00 to 2.40.  Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty. Conversely, simulations with detided time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero) were those for which there was no biological or tidal influence.  While the latter examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regression provided reasonable results given extremes.

Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1,fig:err_surf1}).  Model performance was minimally influenced by magnitude of the tidal advection component ($DO_{adv}$), although performance decreased slightly with increasing tidal effects.  Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2,err_surf2}).  Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.  Scale differences between \cref{fig:err_surf1} and \cref{fig:err_surf2} emphasize that model performance was more affected by characteristics of the \ac{DO} time series rather than the selected window widths.  For example, the range of correlation values comparing the effects of half window widths (averaged across all simulation parameters, \cref{fig:err_surf2}) were approximately half the range of correlations for comparing the effects of simulation parameters (averaged across all half window widths, \cref{fig:err_surf1}).

@@

The weighted regression method was applied to the time series for each station to obtain a detided \ac{DO} time series for estimating metabolism.  Half window widths of six days, one hour, and a tidal proportion of one half were chosen based on a balance between large and small window widths, although we recognize that the chosen widths are somewhat arbitrary and a more exhaustive evaluation should be conducted prior to using the results to inform management actions. Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.  Accordingly, results were evaluated  using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.  Results were also evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.  Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}.  We hypothesized that metabolism esimates using the detided signal would contain less `anomalous' values than those from the observed \ac{DO} time series.  Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values.

Ecosystem metabolism was estimated from the \ac{DO} time series using the open-water technique \citep{Odum56} as described in \citet{Caffrey13}.  The method is used to infer net ecosystem metabolism using the mass balance equation:
\begin{equation} \label{metrate}
\frac{\delta DO}{\delta t} = P - R + D
\end{equation}
where the change in \ac{DO} concentration ($\delta DO$, g O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange at the interface ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}. $D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO}, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.  The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change as the difference between observations at time $t$ and $t+1$.  Areal respiration rates were assumed constant during the night and substracted from daily gross production estimates to yield net ecosystem metabolism (\cref{tab:case_att}).

Detiding had significant effects on the correlations between tidal height changes, \ac{DO} time series, and metabolism estimates (\cref{tab:cor_res}).  Correlations of observed \ac{DO} time series with predicted tidal height were highly significant, with all sites indicating positive relationships, except Padilla Bay where tidal increases were associated with declines in \ac{DO}  concentration.  This suggests that seaward water masses were less anoxic than landward masses, with the opposite being true for Padilla Bay. The detided \ac{DO} time series had greatly reduced correlations with tidal height, although relationships were still significant after detiding likely because of the large sample size for each site (n $\approx$ 17500). Metabolism estimates before and after detiding were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.  Production rates were compared to mean rates of tidal height change during the day, respiration rates were compared to mean rates of change during the night, and net metabolism rates were compared to mean rates of change for the total 24 hour period each day.  Comparison of metabolic rates to tidal changes before and after detiding produced inconsistent results (\cref{tab:cor_res}).  Correlations for Elkhorn Slough and Sapelo Island showed consistent reductions in all three metabolims estimates after detiding.  Correlations for Padilla Bay and Rookery Bay were of opposite sign and greater magnitude after detiding for production and respiration, although net metabolism estimates had reduced correlations.

The percent of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after detiding (\cref{tab:case_res}).  Before detiding, anomalous values ranged from 0.09 (Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration as proportions of the daily estimates. Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (0.13 for production, 0.13 for respiration), and only slightly reduced for Elkhorn Slough (0.17 for production, 0.17 for respiration).  Metabolism estimates using detided \ac{DO} time series had decreased mean production and respiration for Elkhorn Slough, increased mean production and respiration for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Mean net ecosystem metabolism was unchanged for all sites.  Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all cases after detiding.  Graphical results for each case study can be viewed using the link in the \hyperref[multi]{multimedia} section.

An example from Sapelo Island illustrates the effects of weighted regression on \ac{DO} and metabolism estimates (\cref{fig:phase_out,fig:phase_in,fig:case_ex}).   A two-week period in February shows when the tidal changes were both in and out of phase with the diel cycling.  The first week illustrates a period of observation when maximum tide heights are generally out of phase with the diel cycle such that low tides are observed during the middle of the night and the middle of the day (\cref{fig:phase_in}).  The second week illustrates when the maximum tide height occurs during the day and night (\cref{fig:phase_in}).  The effects of tidal height change on the observed \ac{DO} time series are visually apparent in the plots.  Maximum or minimum tidal heights also occur when the contributions of respiration and production are largest (i.e., mid-day or mid-night). The first example illustrates a strong negative bias (less respiration, less production) in the observed \ac{DO} signal from low tides, whereas the second example illustrates a strong positive bias (more respiraiton, more production) in the observed \ac{DO} from high tides. These biases are apparent in the metabolism estimates based on the observed data (\cref{fig:case_ex}).  Anomalous estimates occur when low tides are in phase with the solar cycle, whereas metabolism estimates are likely over-estimated when high tides are in phase with the solar cycle.  The detided time series shows noticeable changes given the direction of bias associated with tidal height and diel period.  \ac{DO} values were higher after detiding when low tides occurred during night and day periods, whereas \ac{DO} values were lower after detiding when high tides occurred during day and night periods.  Changes in metabolism estimates after detiding were also apparent, such that the anomalous values were removed during the first week and the positive bias in the second week is decreased.  Detiding had similar effects for the remaining sites, particularly when tidal changes were in or out of phase with diel periods.

@@

A final point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.  From a management or ecological perspective, the effects of tidal variatoin on daily estimates may not be a primary concern given that seasonal or annual rates may be more relevant for evaluating ecosystem dynamics with continuous monitoring data.  The example from Sapelo Island in the previous section further highlights this point given that mean production and respiration estimates before and after detidng were generally unchanged for the two-week period. \cref{tab:case_res} also indicated that mean annual estimates of production and respiration were unchanged for Rookery Bay and Sapelo Island, whereas production and respiration estimates were significantly different at both mean annual and daily time scales for Elkhorn Slough and Padilla Bay. Although we acknowledge that the specific results may be related to the window widths, this suggests that detiding may contribute to changes in metabolic estimates that are aggregated on longer time periods.  Therefore, an evaluation of the effects of tidal variation on ecosystem metabolism for different periods of observation is critical for understanding practical implications of weighted regression.  Specifically, when should detiding be applied if aggregation of observed data on longer time periods removes potential bias?  A comparison of observed and detided estimates that are aggregated over different periods of observation (e.g., annual, seasonal, monthly) could help address this question.

The observed and detided daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate effects of aggregation on mean production and respiration estimates (\cref{fig:metab_sum1,fig:metab_sum2}).  Mean annual estimates in \cref{tab:case_res} also provided a basis of comparison with monthly and seasonal aggregation. Significant variation in aggregated production and respiration estimates for month and season was observed for each case study.  Detided production and respiration estimates for Padilla Bay and Rookery Bay exhibited seasonal and monthly variation that was more characteristic of expected trends with increases in metabolism during warmer months.  Specifically, production estimates based on observed \ac{DO} were substantially muted for both Padilla Bay (\cref{fig:metab_sum1}) and Rookery Bay (\cref{fig:metab_sum2}) during summer months, whereas values were significantly higher based on the detided data. Results for Sapelo Island suggested that winter and summer months were under- and over-estimated, respectively, based on the observed data.  Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after detiding regardless of the aggregation period.  Overall, these trends emphasize the importance of considering different aggregation periods for interpreting metabolism estimates.  Each case study showed differences in observed and detided values at monthly and seasonal aggregations, whereas only two of the four case studies had mean aggregated estimates that were substantially different (Elkhorn Slough and Padilla Bay, \cref{tab:case_res}).  Periods of observation as long as one year may include significant sources of bias from tidal advection, suggesting the need for applying weighted regression.

@@

Results from the simulations and case studies suggested that weighted regression can be a practical approach for detiding \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.  However, application of the method may only be appropriate under specific situations.  The case studies were chosen based on the expected influence of tidal variaiton given the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.  Despite these similarites among the case studies, detiding had variable effects on metabolism estimates.  The results for Elkhorn Slough and Padilla Bay are particular cause for concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.  Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after detiding suggests a more careful evaluation of the method is needed.  In particular, optimal window widths should be chosen based on the ability to remove tidal effects while preserving the biological signal.

The case studies emphasize a critical challenge in evaluating the effects of physical advection: the true biological signal is not known, or more appropriately, the relative contribution of horizontal advection to bias is not accuratley quantified.  Comparative analyses between systems with varying tidal influence or within-system evaluations of multiple sites at fixed distances are necessary to further validate performance of weighted regression.  Regardless, the current analysis suggests the method is potentially useful and we propose a precautionary approach for its application pending further validation.  Weighted regression should only be applied at macrotidal sites with strong evidence of the confounding effects of tidal advection on biological signals.  A weight-of-evidence approach should be used such that the occurrence of anomalous metabolism estimates, strong correlations between observed \ac{DO} and tide height, and clear visual patterns of tide change on \ac{DO} would suggest detiding is appropriate.  The choice of window widths may also produce varying results.  Optimal widths may be those that indicate a reduction of anomalous estimates, reduction in daily variation of production and respiration, and a reduction of the correlation between \ac{DO} and tide.  Results from window widths that produce large changes in the mean annual estimates should interpreted with caution.  In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation related to production while exhibiting minimal changes with the tide.  Such an approach, combined with further validation, will support informed management decisions through more accurate estimates of ecosystem metabolism.

@@

\caption[Example of detiding a simulated \ac{DO} time series]{Example of detiding a simulated \ac{DO} time series.  Simulated values are those in \cref{fig:do_sim}.  Yellow indicates daylight periods.\label{fig:do_dtd}}
\end{figure}

@@

\caption[Example of creating a simulated \ac{DO} time series]{Example of creating a simulated \ac{DO} time series.  Values were simulated every 30 minutes for 30 days.  Yellow indicates daylight periods.\label{fig:do_sim}}
\end{figure}

@@

\caption[Representative examples of simulated observed \ac{DO} ($DO_{obs}$) and biological \ac{DO} ($DO_{bio})$ as a component of observed) time series created by varying each of four parameters]{Representative examples of simulated observed \ac{DO} ($DO_{obs}$) and biological \ac{DO} ($DO_{bio})$ as a component of observed) time series created by varying each of four parameters: tidal strength of tidal association with \ac{DO} signal using $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.  Parameter values represent the minimum and maximum used in the simulations.\label{fig:sim_ex}}
\end{figure}

@@

\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters: strength of tidal association with \ac{DO} signal $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.  Each tile represents the correlation or error between detided and biological \ac{DO} time series from results for a given combination of simulation parameters.  Results are averaged for all window widths used to evaluate the regressions (\cref{fig:err_surf2}).\label{fig:err_surf1}}
\end{figure}

>>>>

\usepackage{alltt}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{indentfirst}
\usepackage{titlesec}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{outlines}
\usepackage{lineno}
\usepackage{array}
\usepackage{times}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage[position=t]{subfig}
\usepackage{paralist}
\usepackage[noae]{Sweave}
\usepackage{natbib}
\usepackage{array}
\usepackage{pdflscape}
\usepackage{bm}
% \usepackage{showlabels}
\bibpunct{(}{)}{,}{a}{}{,}

@@

Multiple time series were simulated by varying the conditions in \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}. Specifically, the simulated data varied in the relative amount of noise in the measurement ($e_{pro}$, $e_{obs}$), relative amplitude of the diel \ac{DO} component ($DO_{die}$), and degree of association of the tide with the \ac{DO} signal ($DO_{adv}$).  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$. A total of 81 time series were created based on the unique combinations of parameters (\cref{fig:sim_ex}).  Three half window widths for each variable (day, hour of day, and tide height) were evaluated for each simulation: decimal time as 1, 3, and 6 days, time of day as 1, 3, and 6 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the total range given the height at the center of the window.  The window widths were chosen based on preliminary assessments that suggested variation in model performance was adequately captured by these values.  In total, 27 window width combinations were evaluated for each of 81 simulated time series, producing results for 2187 weighted regressions.

The detided or normalized values for each regression were compared to the simulated data to evaluate the ability of weighted regression to reproduce the biological \ac{DO} signals. Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the detided time series and the biological \ac{DO} time series as a known component of the observed.  Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2},   results for each simulation can be viewed using the link in the \hyperref[multi]{multimedia} section).  Mean correlation for all time series and window widths between the detided and biological values was 0.60, with values ranging from -0.78 to 1.00.  Mean error was 1.21, with values ranging from 0 to 2.40.  Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty. Conversely, simulations with detided time series that were identical to the true time series (e.g., correlation of one, \ac{RMSE} of zero) were those for which there was no biological or tidal influence.  Although these examples do not represent real-world scenarios, they were included in the simulations to provide verification that the weighted regression provided reasonable results given extremes.

Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreased performance (\cref{tab:dtd_perf1,fig:err_surf1}).  Model performance was minimally influenced by magnitude of the tidal advection component ($DO_{adv}$), although performance decreased slightly with increasing tidal effects.  Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2,fig:err_surf2}).  Graphical summaries of model performance by simulation parameters (\cref{fig:err_surf1}) and half window widths (\cref{fig:err_surf2}) support the general trends described by \cref{tab:dtd_perf1,tab:dtd_perf2}.  Scale differences between \cref{fig:err_surf1} and \cref{fig:err_surf2} emphasize that model performance was more affected by characteristics of the \ac{DO} time series rather than the selected window widths.  For example, the range of correlation values comparing the effects of half window widths (averaged across all simulation parameters, \cref{fig:err_surf2}) were approximately half the range of correlations for comparing the effects of simulation parameters (averaged across all half window widths, \cref{fig:err_surf1}).

@@

The weighted regression method was applied to the time series for each station to obtain a detided \ac{DO} time series for estimating metabolism.  Half window widths of six days, one hour, and a tidal proportion of one half were chosen based on a balance between large and small window widths, although the chosen window widths are arbitrary and a more exhaustive evaluation should be conducted prior to using the results to inform management. Unlike the simulated data, the true biological \ac{DO} signal was unknown for the case studies.  Accordingly, results were evaluated  using correlations of \ac{DO} and metabolism estimates with tidal height before and after application of the model.  Daily metabolism estimates before and after detiding were compared to the mean rate of tidal height change (i.e., first derivative of the predicted tidal height) for each day during separate solar periods.  Production rates were compared to mean rates of tidal height change during the day, respiration rates were compared to mean rates of change during the night, and net metabolism rates were compared to mean rates of change for the total 24 hour period each day.  Results were also evaluated based on the occurrence of `anomalous' daily production or respiration estimates, where anomalous was defined as negative production during the day and positive respiration estimates during the night.  Anomalous values have been previously attributed to the effects of physical processes on \ac{DO} time series \citep{Caffrey03}.  We hypothesized that metabolism esimates using the detided signal would contain less `anomalous' values than those from the observed \ac{DO} time series.  Although anomalies could be caused by processes other than tidal advection, e.g., abiotic dark oxygen production \citep{Pamatmat97}, we assume that physical processes are the dominant sources of these values given the tidal characteristics at each site.  Finally, means and standard errors of metabolism estimates were evaluated before and after detiding to determine if annual aggregations were significantly different.

Ecosystem metabolism was estimated from the \ac{DO} time series using the open-water technique \citep{Odum56} as described in \citet{Caffrey13}.  The method is used to infer net ecosystem metabolism using the mass balance equation:
\begin{equation} \label{metrate}
\frac{\delta DO}{\delta t} = P - R + D
\end{equation}
where the change in \ac{DO} concentration ($\delta DO$, g O$_2$ m$^{-3}$) over time ($\delta t$, hours) is equal to photosynthetic rate ($P$, g O$_2$ m$^{-3}$ hr$^{-1}$) minus respiration rate ($R$, g O$_2$ m$^{-3}$ hr$^{-1}$), corrected for air-sea gas exchange at the surface ($D$, g O$_2$ m$^{-3}$ hr$^{-1}$) \citep{Caffrey13}. $D$ is estimated as the difference between the \ac{DO} saturation concentration and observed \ac{DO}, multiplied by a volumetric reaeration coefficient, $k_a$ \citep{Thebault08}.  The diffusion-corrected \ac{DO} flux estimates were averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of \ac{DO} change as the difference between observations at time $t$ and $t+1$ on an hourly scale.  Areal respiration rates were assumed constant during the night and substracted from daily gross production estimates to yield net ecosystem metabolism (\cref{tab:case_att}).

Detiding had significant effects on the correlations between tidal height changes, \ac{DO} time series, and metabolism estimates (\cref{tab:cor_res},  see the link in the \hyperref[multi]{multimedia} section for graphical results of each case study).  Correlations of observed \ac{DO} time series with predicted tidal height were highly significant, with all sites indicating positive relationships, except Padilla Bay where tidal increases were associated with declines in \ac{DO}  concentration.  This suggests that seaward water masses were less anoxic than landward masses, with the opposite being true for Padilla Bay. The detided \ac{DO} time series had greatly reduced correlations with tidal height, although relationships were still significant after detiding likely because of the large sample size for each site (n $\approx$ 17500). Comparison of metabolic rates to tidal changes before and after detiding produced inconsistent results (\cref{tab:cor_res}).  Correlations for Elkhorn Slough and Sapelo Island showed consistent reductions in all three metabolims estimates after detiding.  Correlations for Padilla Bay and Rookery Bay were of opposite sign and greater magnitude after detiding for production and respiration, although net metabolism estimates had reduced correlations.

The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after detiding (\cref{tab:case_res}).  Before detiding, anomalous values ranged from 0.09 (Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration. Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (0.13 for production, 0.13 for respiration), and only slightly reduced for Elkhorn Slough (0.17 for production, 0.17 for respiration).  Metabolism estimates using detided \ac{DO} time series had decreased mean production and respiration for Elkhorn Slough, increased mean production and respiration for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Mean net ecosystem metabolism was unchanged for all sites.  Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all cases after detiding.

An example from Sapelo Island illustrates the effects of weighted regression on \ac{DO} and metabolism estimates (\cref{fig:phase_out,fig:phase_in,fig:case_ex}).   A two-week period in February shows when the tidal changes were both in and out of phase with the diel cycling.  Maximum tide heights were generally out of phase with the diel cycle during the first week such that low tides were observed during the middle of the night and the middle of the day (\cref{fig:phase_in}).  The opposite scenario occurred during the second week when the maximum tide height occured during the day and night (\cref{fig:phase_in}).  The effects of tidal height change on the observed \ac{DO} time series are visually apparent in the plots. The first week illustrates a strong negative bias (less respiration, less production) in the observed \ac{DO} signal from low tides at mid-day and mid-night, whereas the second example illustrates a strong positive bias (more respiration, more production) in the observed \ac{DO} from high tides. These biases are apparent in the metabolism estimates using the observed data (\cref{fig:case_ex}).  Anomalous estimates occur when low tides are in phase with the solar cycle, whereas metabolism estimates are likely over-estimated when high tides are in phase with the solar cycle.  The detided time series shows noticeable changes given the direction of bias from the phasing between tidal height and diel period.  \ac{DO} values were higher after detiding when low tides occurred during night and day periods, whereas \ac{DO} values were lower after detiding when high tides occurred during day and night periods (\cref{fig:phase_out,fig:phase_in}).  Changes in metabolism estimates after detiding were also apparent, such that the anomalous values were removed during the first week and the positive bias in the second week is decreased.  Detiding had similar effects for the remaining sites, particularly when tidal changes were strongly in or out of phase with diel periods.

@@

A final point of concern is the period of observation within which observed \ac{DO} is affected by tidal height changes and the extent to which this affects the interpretation of ecosystem metabolism.  From a management or ecological perspective, the effects of tidal variation on daily estimates may not be a primary concern given that seasonal or annual rates may be more relevant for evaluating ecosystem dynamics with continuous monitoring data.  The example from Sapelo Island in the previous section highlights this point given that mean production and respiration estimates before and after detiding were generally unchanged for the two-week period. \cref{tab:case_res} also indicated that mean annual estimates of production and respiration were unchanged for Rookery Bay and Sapelo Island, whereas production and respiration estimates were significantly different for Elkhorn Slough and Padilla Bay. Although we acknowledge that the specific results may be related to the window widths, this suggests that tidal effects contribute to changes in metabolic estimates from observed time series that are aggregated on longer time periods.  Therefore, an evaluation of the effects of tidal variation on ecosystem metabolism for different periods of observation is critical for understanding practical implications of weighted regression.  Specifically, when should detiding be applied if aggregation of observed data on longer time periods removes potential bias?  A comparison of observed and detided estimates that are aggregated over different periods of observation (e.g., annual, seasonal, monthly) could help address this question.

The observed and detided daily estimates were averaged by month and season (Fall, Spring, Summer, and Winter) for each case study to evaluate effects of aggregation on mean production and respiration estimates.  Mean annual estimates in \cref{tab:case_res} also provided a basis of comparison with monthly and seasonal aggregation. Significant variation in aggregated production and respiration estimates for month and season was observed for each case study (\cref{fig:metab_sum1,fig:metab_sum2}).  Detided production and respiration estimates for Padilla Bay and Rookery Bay exhibited seasonal and monthly variation that was more characteristic of expected trends with increases in metabolism during warmer months.  Specifically, production estimates based on observed \ac{DO} were substantially muted for both Padilla Bay (\cref{fig:metab_sum1}) and Rookery Bay (\cref{fig:metab_sum2}) during summer months, whereas values were significantly higher based on the detided data. Results for Sapelo Island suggested that winter and summer months were under- and over-estimated, respectively, based on the observed data.  Results for Elkhorn Slough varied significantly such that production and respiration were significantly reduced after detiding regardless of the aggregation period.  Overall, these trends emphasize the importance of considering different aggregation periods for interpreting metabolism estimates.  Each case study showed differences in observed and detided values at monthly and seasonal aggregations, whereas only two of the four case studies had mean aggregated estimates that were substantially different (Elkhorn Slough and Padilla Bay, \cref{tab:case_res}).  Periods of observation as long as one year may include significant sources of bias from tidal advection, suggesting the need for applying weighted regression given careful consideration of appropriate window widths.

@@

Results from the simulations and case studies suggested that weighted regression can be a practical approach for detiding \ac{DO} time series to remove the effects of physical advection on estimates of ecosystem metabolism.  However, application of the method may only be appropriate under specific situations.  The case studies were chosen based on the relatively high proportion of metabolism estimates that were anomalous and the strength of correlation between the observed \ac{DO} time series and tidal height.  Despite these similarites among the case studies, detiding had variable effects on metabolism estimates.  The results for Elkhorn Slough and Padilla Bay are of particular concern given that mean annual estimates were substantially different compared to those from the observed \ac{DO} time series.  Although the correlation of \ac{DO} and tidal height was reduced for both cases, in addition to a reduction of anomalous estimates, the relative change in mean metabolism before and after detiding suggests a more careful evaluation of the method is needed.  In particular, alternative window widths should evaluated for the ability to remove tidal effects while preserving the biological signal.  The window widths in the above analysis may have removed variation in the \ac{DO} signal from both of these sources.

Although the above analyses suggest the approach has merit, the case studies emphasize a critical challenge in applying weighted regression to monitoring data. Specifically, the true biological signal is not known and the relative contribution of horizontal advection to bias is not accurately quantified with the available data.  Comparative analyses between systems with varying tidal influence or within-system evaluations of multiple sites at fixed distances are necessary to further validate the performance of weighted regression.  In the absence of additional validation, we propose a precautionary approach for application of the weighted regression to monitoring data.  Weighted regression may be most effective at macrotidal sites with strong evidence of the effects of tidal advection on biological signals.  A weight-of-evidence approach should be used such that the occurrence of anomalous metabolism estimates, strong correlations between observed \ac{DO} and tide height, and clear visual patterns of tide change on \ac{DO} would suggest detiding is appropriate.  The choice of window widths may also produce varying results.  Window widths that produce large changes in mean annual estimates should be interpreted with caution.  In general, a pragmatic approach is emphasized such that results should be evaluated based on the preservation of diel variation from production while exhibiting minimal changes with the tide.  Such an approach, combined with further validation, will support informed management decisions through more accurate estimates of ecosystem metabolism.

@@

\caption[Example of detiding a simulated \ac{DO} time series]{Example of detiding a simulated \ac{DO} time series.  $DO_{obs}$ represents an additive time series as in \cref{do_obs_all}, $DO_{tid}$ represents the predicted values of \ac{DO} conditional on tidal height and time (\cref{do_tid}), and $DO_{nrm}$ represents the detided values of \ac{DO} conditional on constant tidal height and time (\cref{do_nrm}).  Yellow indicates daylight periods.\label{fig:do_dtd}}
\end{figure}

@@

\caption[Example of each component of a simulated \ac{DO} time series for testing weighted regression]{Example of each component of a simulated \ac{DO} time series for testing weighted regression.  The time series were created using \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}. Yellow indicates daylight periods.\label{fig:do_sim}}
\end{figure}

@@

\caption[Representative examples of simulated time series of observed \ac{DO} ($DO_{obs}$) and biological \ac{DO} ($DO_{bio}$, as a component of observed) created by varying each of four parameters]{Representative examples of simulated time series of observed \ac{DO} ($DO_{obs}$) and biological \ac{DO} ($DO_{bio}$, as a component of observed) created by varying each of four parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\epsilon_{pro}$), amount of observation observation uncertainty ($\epsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).  Parameter values represent the minimum and maximum used in the simulations as mg L$^{-1}$ of \ac{DO}.\label{fig:sim_ex}}
\end{figure}

@@

\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series ($DO_{dtd}$) from weighted regression with `true' biological \ac{DO} ($DO_{bio}$) for varying simulation parameters: strength of tidal association with \ac{DO} signal ($DO_{adv}$), amount of process uncertainty ($\epsilon_{pro}$), amount of observation observation uncertainty ($\epsilon_{obs}$), and strength of diel \ac{DO} component ($DO_{die}$).  Each tile represents the correlation or error from results for a given combination of simulation parameters averaged for all window widths (\cref{fig:err_surf2}).\label{fig:err_surf1}}
\end{figure}
